Formation of porous crystals via viscoelastic phase separation

### 2017-01-26 nature/RebuttalLetter.tex

\noindent
\begin{Large}
{\bf Response to Referee \#3}
\end{Large}

\vskip 0.5cm

We would like to thank the Referee for his/her comments. In the following we respond to the new points raised.

\begin{referee}
\noindent
1) I was aware of the different frame rates on the two movies but I agree that the time stamps are helpful. Figure S5 addresses my concern about reproducibility at the single state point, and the different geometry itself is helpful too. I think the authors should be up front in the manuscript that the mechanism discussed is found in a very specific window (from the figures and SI, I think it is $\phi\sim 0.33$ and $c_p \sim0.39$ (plus minus) 0.01 mg/g) by moving this point into the abstract.
\end{referee}

We are glad to hear that the reviewer found that the time stamps are helpful and also Fig. S5 addresses his/her concern.      
Following the Referee's recommendation, we have put the following sentence in the abstract: ``We clarify the physical requirements necessary for this to happen.'' Although this does not explicitly specify the points we observed crystal gels, our new analysis (see below) 
has revealed the physical conditions required for observing crystal-gel formation. We have also discussed this issue in our revised manuscript with the new analysis shown in a new Fig. 2. We stress that the conditions for crystal-gel formation are not stringent and well within the experimental accessible states, as we previously demonstrated in our reproducibility tests inside a capillary. Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei. To our knowledge, this is the first experimental demonstration of this process: it is also not easy to be captured by computer simulations, as they often neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition. The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.

@@

\begin{referee}
2) The addition of the SI movie 3 and the discussion of Figure 2b addresses this concern for me. It's the percolation that is critical to the authors claim of a crystal gel - the percolated structure needs to persist over long times, otherwise, for example, the gel wouldn't have a finite low frequency elastic modulus. I think SI move 3 does show this.
\end{referee}

We are glad to hear that the reviewer thought that SI movie 3 and the discussion of Fig. 2b have addressed his/her concern and supported our claim of a crystal gel.

@@

\begin{referee}
4) Figure S4 is critical to the interpretation, including information about all the transition pathways.
\end{referee}

We are glad to know that the reviewer thought that Fig. S4 is useful for supporting the interpretation.

@@

\begin{referee}
 5) The revised Figure 4, along with Figure S4, are very helpful to resolving the relative contributions of the three crystalline pathways. In Figure 4b,c, I still think that surface particles should somehow be distinguished - they represent a distinct state and a kind of consistency check, because the different crystallization pathways should all be mediated by this state. Without including the surface states in the primary figure, I don't see how the reader can inspect Figure S4, and get to the 54/34/11 abundances for direct, Bergeron and Ostwald processes. These abundances are the crucial claim of the paper.
 \end{referee}

We are glad to hear that the revised Fig. 4, along with Fig. S4, are very useful to resolving the relative contributions of the three crystalline pathways. Following the kind suggestion of the reviewer, we have now included the fraction of surface particles in Fig. 6b and added the related discussion.

@@

\begin{referee}
In summary the confirmatory experiments (Fig S5 and Movie S3), the additional analysis of the crystalline state (Fig 3S) and the full transition analysis (rev Fig 4 and F S4) offer significant new support of the authors' claims; the authors have something new relative to prior literature. At this point, I think that the authors should: (i) better acknowledge that what is observed here is found in a quite specific region of the state diagram ($phi\sim 0.33$ and $c_p\sim 0.39$ {plus minus} 0.01 mg/g); (ii) better connect Figure 4 and Figure S4 by accounting for surface states in the former figure.

\end{referee}

We are glad to hear that the Referee thought that the previous revisions offered significant new support of the authors' claims. 
On the comment (i), as we explained above, we have clarified the physical conditions required for crystal-gel formation to take place. We have also discussed how we can widen the region where this is to be observed (see page 14). 
(ii)  As mentioned above, following the suggestion, in the new version of the manuscript we follow the Referee's recommendation and include in Fig. 6b the fraction of surface particles, and update the description accordingly.
